Scaling up the hyperparameter grid for Naomi

### 2023-03-07 src/dev_scale-grid/scale-grid.Rmd
Updating PCA approach for creating a grid

Such that `C` can be obtained by `E %*% diag(lambda) %*% t(E)`:

@@

```{r}
create_grid <- function(d, k) {
  gg <- mvQuad::createNIGrid(dim = d, type = "GHe", level = k)
  list(nodes = mvQuad::getNodes(gg), weights = as.numeric(mvQuad::getWeights(gg)))
}

gg_full <- create_grid(d = length(hypers), k = 2)
```

Start with `p = 5` dimensions kept of the PCA.
Then create a dense AGHQ grid with `k = 2` on 5 dimensions.

```{r}
p <- 5
d <- dim(E)[1]
gg_p <- create_grid(d = p, k = 2)
```

@@

```{r}
(n_nodes <- nrow(gg_p$nodes))
```

Extend this into a 32 x 24 sparse matrix with zeros in the other locations:

```{r}
# i and j specify the non-zero locations in the matrix
(nodes_p <- sparseMatrix(
  i = rep(1:n_nodes, times = p),
  j = rep(1:p, each = n_nodes),
  x = as.numeric(gg_p$nodes),
  dims = c(n_nodes, d)
))
```

How does the reconstruction of the covariance matrix with `r p` components look?

```{r}
E_p <- E[, 1:p]
Lambda_p <- Lambda[1:p, 1:p]
C_p <- E_p %*% Lambda_p %*% t(E_p)

@@

C_p_plot <- reshape2::melt(as.matrix(C_p)) %>%
  ggplot(aes(x = Var1, y = Var2, fill = value)) +
  labs(x = "", y = "", fill = "Cp[i, j]") +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(x = "i", y = "j") +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank()
  )

C_plot + C_p_plot
```

Want to do $E_p \Lambda_p^{1/2}$ multiplied by the nodes:

```{r}
E_p <- sparseMatrix(i = rep(1:p, times = p), j = rep(1:p, each = p), x = as.numeric(E[1:p, 1:p]), dims = c(d, d))
Lambda_p <- sparseMatrix(i = 1:p, 1:p, x = diag(Lambda)[1:p], dims = c(d, d))

X_p <- E_p %*% sqrt(Lambda_p)

new_nodes <- X_p %*% t(nodes_p)

t(new_nodes)
```

```{r}
# pca_rescale_grid <- function() {
# 
# }
```

```{r}
# gg5 <- pca_rescale_grid(p = 5)
# gg7 <- pca_rescale_grid(p = 7)
# gg7 <- pca_rescale_grid(p = 9)
```

>>>>

Such that `C` can be obtained by $E \Lambda E^\top$, or in code `E %*% diag(lambda) %*% t(E)`:

@@

```{r}
create_grid <- function(d, k) {
  gg <- mvQuad::createNIGrid(dim = d, type = "GHe", level = k)
  list(nodes = mvQuad::getNodes(gg), weights = as.numeric(mvQuad::getWeights(gg)))
}
```

Start with `s = 5` dimensions kept of the PCA.
Then create a dense AGHQ grid with `k = 2` on 5 dimensions.

```{r}
s <- 5
d <- dim(E)[1]
gg_s <- create_grid(d = s, k = 2)
```

@@

```{r}
(n_nodes <- nrow(gg_s$nodes))
```

How does the reconstruction of the covariance matrix with `r s` components look?

```{r}
E_s <- E[, 1:s]
Lambda_s <- Lambda[1:s, 1:s]
C_s <- E_s %*% Lambda_s %*% t(E_s)

@@

C_s_plot <- reshape2::melt(as.matrix(C_s)) %>%
  ggplot(aes(x = Var1, y = Var2, fill = value)) +
  labs(x = "", y = "", fill = "C_s[i, j]") +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(x = "i", y = "j") +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank()
  )

C_plot + C_s_plot
```

Let $a \in \mathbb{R}^d$ and $b \in \mathbb{R}^s$.
The PCA provides a mapping $f: \mathbb{R}^d \to \mathbb{R}^s$ which compresses each $a$ to a smaller $b$ given by $f(a) = (E_s \Lambda_s^{1/2})^\top a = b$.
Then $f^{-1}$ is given by $f^{-1}(b) = E_s \Lambda_s^{-1/2} b = a$, using the fact that $E_s^\top = E_s^{-1}$.
So we need to multiply our nodes by $E_s \Lambda_s^{- 1/2}$ as follows:

```{r}
a_nodes <- gg_s$nodes
b_nodes <- t(E_s %*% diag(lambda[1:s]^{-0.5}) %*% t(a_nodes))

dim(a_nodes)
dim(b_nodes)
```

Add on the modes:

```{r}
for(j in 1:d) b_nodes[, j] <- b_nodes[, j] + m[j]
```

Adjust the weights:

```{r}
gg_s$weights
det(Lambda)
det(Lambda_s)
```

```{r}
#' @param m Mean vector
#' @param C Covariance matrix
#' @param s Small grid dimension
#' @param k Number of points per small grid dimension
pca_rescale_grid <- function(m, C, s, k) {
  lambda <- eigenC$values
  Lambda <- diag(lambda)
  E <- eigenC$vectors
  E_s <- E[, 1:s] 
  gg_s <- create_grid(d = s, k = k)
  a_nodes <- gg_s$nodes
  b_nodes <- t(E_s %*% diag(lambda[1:s]^{-0.5}) %*% t(a_nodes))
  for(j in 1:d) b_nodes[, j] <- b_nodes[, j] + m[j]
  b_weights <- gg_s$weights * sqrt(det(Lambda))
  return(list(nodes = b_nodes, weights = b_weights))
}
```

```{r}
gg5 <- pca_rescale_grid(m = m, C = C, s = 5, k = 2)
gg7 <- pca_rescale_grid(m = m, C = C, s = 7, k = 2)
gg9 <- pca_rescale_grid(m = m, C = C, s = 9, k = 2)
```
